---
title: "NCED: Part 1"
date: "November 22, 2019"
output: html_document
---

```{r library}
library(knitr)
library(tidyverse)
library(formattable)
library(tidyverse)
library(skimr)
```

##Project Assignment #1					Due: Tuesday, October 8 by 3:00PM
Value: 20% of your project grade (4% of your final grade)
1. Describe your topic. What relationship/phenomenon do you want to study? This should be a research question that has been refined and polished since your initial discussions with me on your topic.

I hypothesize that the more development that a region has, the more people will want to live there, and the more people that will need to support the industries that are in it. An area with more people, a better educated workforce, and more people and ideas being imported will have better economic prospects. Businesses are attracted to a large, well educated labor force, and well educated workers are attracted to places that have the services they need to support their quality of life.
Citing Lucas 2015; Manuelli and Seshadri 2014; Gennaloli et al. 2013; Hendricks 2002, Fan et. al. find that human capital is a major factor in economic growth.

I want to study the relationship between local geographic area and economic outcome. I have to focus on differences between each geographic area. For this regression, I will focus on population because human capital has an empirical relationship to economic development. In this study, economic development will be examined primarily through Median Per Capita Income, as well as homeownership and employment variables. 

also have data on other development indicators that I think may have an interfering effect on the population. I tried to gather variables that address the composition of the population as well, including race, number of people who lived outside of the county 5 years ago, etc. to study the influences of other factors. Optimally, when controlling for these factors, the population effect will still be significant due to the sheer amount of labor and industry required to support higher paying jobs for the workforce to participate in. 
(Total population)/income relationship
North Carolina counties

2. Why is this topic interesting and important? What policy makers or ﬁrms would be interested in the topic?

Should there be a relationship, it could be an indicator about poverty and economic development. This could develop into a study about the development of cities and towns as it relates to economic development, a question that would have significant impacts on how we ensure that people have fair access to resources.

3. What is the population you are studying? What sample are you using? How representative is your sample of the population? What problems might there be with your sample?

I’ll be studying the entire population of North Carolina (or the populations of every Metropolitan area in the South). I am using a sample of every county in North Carolina. My sample should be relatively accurate to the population because it is population data (I think) collected by each county in North Carolina. All of my data is from 2015 except for data from the Count table, which is unidentified by a ‘Year’ variable. This may pose a problem as to the validity of my findings, due to the fact that it may not be accurate to the year 2015. My sample may have some issues with assignment error, as I did have to transform some of the columns and rows to have clean data. 
4. What is the source of your data? Did you have transform the data to get them into their present form? If so, what did you do?
My sample will be coming from the North Carolina Office of State Budget and Management (OSBM) in the Log Into North Carolina (LINC) dataset. I did have to join data from separate surveys to make sure that I could get the information I needed. It was an inner join on Area Name. This included data on housing, population, income, and education.
```{r}
## Data Load ##
census = read_delim(
  "LINC/census-population-and-housing-linc.csv",
  delim = ";")
count = read_csv(
  "LINC/ncccc-nc-complete-count-committee.csv")
education = read_delim(
  "LINC/education.csv",
  delim = ";")
employment = read_delim(
  "LINC/employment-and-income-linc.csv",
  delim = ";")

## format raw col as rows ##
census = census %>% 
  pivot_wider(names_from = Variable,values_from = Value) %>% 
  select(-Year) %>% 
  na.omit() #removes entries with null values

count = count %>% 
  pivot_wider(names_from = Race,values_from = value) %>% 
  select(-main_facet)

education = education %>% 
  pivot_wider(names_from = Variable,values_from = Value) %>% 
  select(-`Area Type`)

employment = employment %>% 
  pivot_wider(names_from = Variable,values_from = Value) %>% 
  select(-Year)

#### join datasets together ####
# cc_table = inner_join(census1,count1,by='Area Name') # join edu and count
# ee_table = inner_join(education1,employment1,by='Area Name')
# full = inner_join(cc_table,ee_table,by='Area Name')

full = inner_join(census,count,by='Area Name') %>% 
  inner_join(education,by='Area Name') %>% 
  inner_join(employment,by='Area Name') %>% 
  mutate(maj_white = as.logical(ifelse(`White alone`>= 67.0,1,0))) %>% 
  # condense all the race variables into one column
  nest(race = c(6:13)) %>% 
  # separate latitude/longitude
  separate(geo_point_2d, sep = ",", into = c("lat","long"),remove = TRUE) %>% 
  mutate(lat = as.numeric(lat),
         long = as.numeric(long)) %>% 
  nest(coordinates = c(lat,long)) %>% 
  select(-Year) %>% 
  # better variable names
  rename(area = 1,
    medi_owner = 2,
    outsiders = 4,
    population = 3,
    sat = 5,
    avg_employ = 8,
    avg_wage = 10,
    employ_resid = 9,
    fam_income = 11,
    college = 6,
    high_sch = 7)
```
I am also exploring county level data from the BEA.

5. Describe the characteristics of each variable: units of measurement, whether it is categorical or quantitative, discrete or continuous. If a variable is categorical, provide the coding scheme used to convert it to a dummy variable.
```{r}
colnames(full)
```

From the Census Population and Housing (LINC) data: Population Estimate (BEA per Capita Denominator) [persons, quant, cont]; Persons Living Outside This County Five Years Ago [persons (prop), quant, cont]; Median Value of Owner Occupied Units [$, quant, cont]
From NC Complete Count Committee data: American Indian and Alaska Native alone [%, cat, cont]; Asian alone [%, cat, cont]; Black or African American alone [%, cat, cont]; Hispanic or Latino (of any race) [%, cat, cont]; Native Hawaiian and Pacific Islander alone [%, cat, cont]; Some Other Race alone [%, cat, cont]; Two or more races [%, cat, cont]; White alone [%, cat, cont]
With a race variable, I coded it as being majority white or not. To determine this scheme, I used the average of the white population variable from NC State over the fall and spring semesters in 2015. F: 66.4%; P: 67.7%; Avg: 67.05%	if the white population is >= 67.05 then 1, if <= then 0.
From Employment and Income data: Estimated Median Family Income(HUD) [\$, quant, cont]; Employment by Place of Residence [?, quant, cont]; Average Annual Wage per Worker [\$, quant, cont]; Average Annual Employment by Place of Work [?, quant, cont]
From Education: SAT Grand Total Average Score [none, quant, cont]; College Graduates Age 25 Up [persons, quant, cont]; High School Graduates Age 25 Up [persons, quant, cont];
SECOND DUMMY: EAST WEST DIVIDE
When plotting a number of development variables, there were similar patterns throughout the data where certain counties were consistetly significantly outliers. This is an interesting point of examination, because 

6. What is your dependent variable? Which of your independent variables is the primary focus of the project? (If you do not have a primary independent variable, which independent variable do you suspect will play the largest role in explaining your dependent variable?) Why did you include the other independent variables (these ones that are not the primary focus)? What factors are these variables controlling for?

The dependent variable I'm using in my sample is `fam_income`, the Estimated Median Family Income(HUD) by county. This variable is an aggregate measure of the financial prosperity of people living in each county. I considered using another income measure, like wage or employment, but median household income is a better way of understanding the overall finances of the individual person. 

I focus heavily on population indicators, movement, and education. I think these three variables will play the largest role in determining the hotspots of NC economic success. ‘Population Estimate’ should paint a good picture of whether or not there is a city or a large economic pull factor in a county. This indicator coupled with ‘Persons Living Outside This County Five Years Ago’ will complete the picture of large economic hubs, identifying areas of c new job creation and higher wages. 
As for the remaining variables I decided to include in the dataset, I chose them because they could all be indicators of economic development in an area, and I would like to control for their effect when I run my OLS regression.

7. Are there variables that you would like to include but do not have data for? How might those omissions affect your results?
In 1983, North Carolina put together a commission to study and improve regional development. The program focuses on industry clusters, and has been wildly successful in developing major metropolitan regions, such as the Research triangle Park and the greater Charlotte metro. The regions include:
*AdvantageWest
*Charlotte Region
*Eastern Region
*Northeast Region
*Southeast Region
*Piedmont Triad
*Research Triangle
```{r region}
INTERACTIVE MAP
Alamance
Piedmont Triad (Central) Prosperity Zone

Alexander
Northwestern Prosperity Zone

Alleghany
Northwestern Prosperity Zone

Anson
Southwestern Prosperity Zone

Ashe
Northwestern Prosperity Zone

Avery
Northwestern Prosperity Zone

Beaufort
Northeast Prosperity Zone

Bertie
Northeast Prosperity Zone

Bladen
Sandhills (South Central) Prosperity Zone

Brunswick
Southeast Prosperity Zone

Buncombe
Western Prosperity Zone

Burke
Northwestern Prosperity Zone

Cabarrus
Southwestern Prosperity Zone

Caldwell
Northwestern Prosperity Zone

Camden
Northeast Prosperity Zone

Carteret
Southeast Prosperity Zone

Caswell
Piedmont Triad (Central) Prosperity Zone

Catawba
Northwestern Prosperity Zone

Chatham
North Central Prosperity Zone

Cherokee
Western Prosperity Zone

Chowan
Northeast Prosperity Zone

Clay
Western Prosperity Zone

Cleveland
Southwestern Prosperity Zone

Columbus
Sandhills (South Central) Prosperity Zone

Craven
Southeast Prosperity Zone

Cumberland
Sandhills (South Central) Prosperity Zone

Currituck
Northeast Prosperity Zone

Dare
Northeast Prosperity Zone

Davidson
Piedmont Triad (Central) Prosperity Zone

Davie
Piedmont Triad (Central) Prosperity Zone

Duplin
Southeast Prosperity Zone

Durham
North Central Prosperity Zone

Edgecombe
North Central Prosperity Zone

Forsyth
Piedmont Triad (Central) Prosperity Zone

Franklin
North Central Prosperity Zone

Gaston
Southwestern Prosperity Zone

Gates
Northeast Prosperity Zone

Graham
Western Prosperity Zone

Granville
North Central Prosperity Zone

Greene
Southeast Prosperity Zone

Guilford
Piedmont Triad (Central) Prosperity Zone

Halifax
Northeast Prosperity Zone

Harnett
North Central Prosperity Zone

Haywood
Western Prosperity Zone

Henderson
Western Prosperity Zone

Hertford
Northeast Prosperity Zone

Hoke
Sandhills (South Central) Prosperity Zone

Hyde
Northeast Prosperity Zone

Iredell
Southwestern Prosperity Zone

Jackson
Western Prosperity Zone

Johnston
North Central Prosperity Zone

Jones
Southeast Prosperity Zone

Lee
North Central Prosperity Zone

Lenoir
Southeast Prosperity Zone

Lincoln
Southwestern Prosperity Zone

Macon
Western Prosperity Zone

Madison
Western Prosperity Zone

Martin
Northeast Prosperity Zone

McDowell
Northwestern Prosperity Zone

Mecklenburg
Southwestern Prosperity Zone

Mitchell
Northwestern Prosperity Zone

Montgomery
Sandhills (South Central) Prosperity Zone

Moore
Sandhills (South Central) Prosperity Zone

Nash
North Central Prosperity Zone

New Hanover
Southeast Prosperity Zone

Northampton
Northeast Prosperity Zone

Onslow
Southeast Prosperity Zone

Orange
North Central Prosperity Zone

Pamlico
Southeast Prosperity Zone

Pasquotank
Northeast Prosperity Zone

Pender
Southeast Prosperity Zone

Perquimans
Northeast Prosperity Zone

Person
North Central Prosperity Zone

Pitt
Northeast Prosperity Zone

Polk
Western Prosperity Zone

Randolph
Piedmont Triad (Central) Prosperity Zone

Richmond
Sandhills (South Central) Prosperity Zone

Robeson
Sandhills (South Central) Prosperity Zone

Rockingham
Piedmont Triad (Central) Prosperity Zone

Rowan
Southwestern Prosperity Zone

Rutherford
Western Prosperity Zone

Sampson
Sandhills (South Central) Prosperity Zone

Scotland
Sandhills (South Central) Prosperity Zone

Stanly
Southwestern Prosperity Zone

Stokes
Piedmont Triad (Central) Prosperity Zone

Surry
Piedmont Triad (Central) Prosperity Zone

Swain
Western Prosperity Zone

Transylvania
Western Prosperity Zone

Tyrrell
Northeast Prosperity Zone

Union
Southwestern Prosperity Zone

Vance
North Central Prosperity Zone

Wake
North Central Prosperity Zone

Warren
North Central Prosperity Zone

Washington
Northeast Prosperity Zone

Watauga
Northwestern Prosperity Zone

Wayne
Southeast Prosperity Zone

Wilkes
Northwestern Prosperity Zone

Wilson
North Central Prosperity Zone

Yadkin
Piedmont Triad (Central) Prosperity Zone

Yancey
Northwestern Prosperity Zon
```
Having this data would have made it easier to see the direct effect of stimulus in NC counties and regions, as they each have targeted development plans. This is a major ommitted variable that will certainly bias my results. It will be difficult to see the complete picture of growth.
I also wanted to include the labor force participation rate, which may have elucidated whether or not each county has a sufficient number of people to support these industries, and labor force participation by demographic group to see if it is dominated by people with or without college degrees, women, immigrants to the county and minorities.

8. Print the first page of your data in the Excel spreadsheet or gretl table. Do not print or submit more than one page because doing so is wasteful. This requirement assures me that you have already collected the data.
```{r head echo=FALSE}
head(full,15)
```